Synthesising and Verifying Closed-loop Digital Control Systems

### 2016-09-23 paper/paper.tex
error bound verification
add Iury's text related to UNCERTAINTY MODEL FOR QUANTIZATION ERROR

@@

@@

>>>>

\newtheorem{myassumption}{Assumption}
\newtheorem{mylemma}{Lemma}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A Uncertainty Model for Quantization Error} 
\label{sec:uncertainty-model-quantization-error}

In this study, an uncertain plant model $\hat{G}(z)$ is split into two terms:
%
\begin{equation}
\label{eq:uncertainplant}
\hat{G}(z)=G(z)+\Delta{G(z)},
\end{equation}

\noindent where $G(z)$ is the discretized nominal plant model and $\Delta{G(z)}$ is the additive uncertainty, which is related to parametric uncertainties~\cite{Bessa16}. Here, an uncertainty model to accommodate the quantization noise effects on $\hat{G}(z)$ is proposed. As a result, Eq.~\eqref{eq:uncertainplant} is rewritten as
%
\begin{equation}
\label{eq:uncertainFWLplant}
\hat{G}(z)=G(z)+\Delta{G_{Q}(z)}+\Delta{G_{P}(z)},
\end{equation}

\noindent where $\Delta{G_{Q}(z)}$ represents the quantization uncertainty model and $\Delta{G_{P}(z)}$ describes the parametric uncertainty model. Thus, a model for $\Delta{G_{Q}(z)}$ that can represent the quantization noise effect in the plant output, and consequently the closed-loop system output, is proposed. For convenience, the parametric uncertainties will be despised here.

Consider the closed-loop sampled system in Figure~\ref{fig:sampledsystem}, where $C(z)$ is the digital controller model, whose control output signal is quantized ($Q2$) and interpolated via a ZOH using digital-to-analog converter (DAC). $G(s)$ is the continuous-time model of the plant, from which output samples are taken and quantized ($Q1$) by means of an analog-to-digital converter (ADC) synchronized with the DAC. Each quantizer ($Q1$ and $Q2$) will present some contribution for the total quantization noise.
%
\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{figures/hsystembd.pdf}
\caption{Sampled system.}
\label{fig:sampledsystem}
\end{figure}

First, some assumptions have to be established to obtain a model for $\Delta{G_{Q}(z)}$.
%
\begin{myassumption}
\label{whitenoise}
The quantization noises $\nu_{1}$ (from the $Q1$) and $\nu_{2}$ (from the $Q2$) are white noises with zero mean.
\end{myassumption}

\begin{myassumption}
The quantization noises $\nu_{1}$ and $\nu_{2}$ are uncorrelated with each other.
\end{myassumption}

\begin{myassumption}
The DAC and ADC are synchronized, {\it i.e.}, there is no delay between sampling the plant output at ADC and updating the DAC accordingly.
\end{myassumption}

\begin{myassumption}
The controller and plant are linear. In particular, the digital controller does not contain nonlinearities such as overflows, saturation, and limit-cycle oscillations.
\end{myassumption}

\begin{myassumption}
The DAC interpolator is an ideal ZOH; it takes the discrete value of control signal $u(k)$ at instant $k$ and produces a rectangular pulse of height $u(k)$ and width equal to the sample time $T$.
\end{myassumption}

Given those assumptions, the present approach is further based on the following lemmas.
%
\begin{mylemma}
\cite{widrow1956} The second moment (variance) of the PDF of quantization noise is given by \textcolor{red}{expand the acronym PDF}
\begin{equation}
\label{eq:variancelemma}
\sigma_{Q}^{2}=\frac{q^{2}}{12},
\end{equation}
\end{mylemma}

\noindent where $q$ is the step size of quantizer.

\begin{mylemma} \cite{astrom1997computer}
The pulse transfer function $G(z)$ is equivalent at the sample times to the continuous transfer function $G(s)$ driven by ZOH. This pulse transfer function can be computed as
\begin{equation}
\label{eq:pulsetf}
G(z)=(1-z^{-1})\mathcal{Z}\left\lbrace{\mathcal{L}^{-1}\left\lbrace{\frac{G(s)}{s}}\right\rbrace_{t=kT}}\right\rbrace.
\end{equation}
\end{mylemma}

Considering the quantization noise $\nu(k)$ as an output error in the discrete plant model, {\it i.e.}, it does not affect the plant dynamics, but its output samples, it is easy to prove that $\Delta{G_{Q}(z)}$ can be obtained from
%
\begin{equation}
\label{eq:quantization_tf}
\Delta{G_{Q}(z)}=\frac{\nu(z)}{U(z)},
\end{equation}

\noindent where $\nu(z)$ is the z-transform of the random process $\nu(k)$, which can be calculated through the power spectral density of $\nu(k)$~\cite{poularikas2000transforms}. Thus, the $\Delta{G_{Q}(z)}$ can expressed as
%
\begin{equation}
\label{eq:deltag_var}
\Delta{G_{Q}(z)}=\frac{\sigma_{Q}^{2}}{U(z)},
\end{equation}

\noindent where $\sigma_{Q}^{2}$ is the variance of quantization noise and $U(z)$ is the z-transform of control signal ({\it i.e.}, the z-transform of the signal from the digital controller), which is provided to the plant by means of ZOH.
%
\begin{proof}
Suppose that the plant discrete model in Figure~\ref{fig:sampledsystem} can be represented by the following pulse transfer function (without quantization noise effect): 
\begin{equation}
\label{eq:tfwithout}
G(z)=\frac{Y(z)}{U(z)}=\frac{B(z)}{A(z)}=\frac{b_{0}+b_{1}z^{-1}+b_{2}z^{-2}+\cdots+b_{M}z^{-M}}{1+a_{1}z^{-1}+a_{2}z^{-2}+\cdots+a_{N}z^{-N}},
\end{equation}
\noindent where $Y(z)$ is the z-transform of the plant output signal without noise $y(k)$, $U(z)$ is the z-transform of the control signal, which is provided to the plant, $A(z)$ and $B(z)$ are the plant pulse transfer function numerator and denominator, respectively.

Considering the quantization noise $\nu(k)$ as an output error in the discrete plant model, an additive noise in the plant output $y(k)$ resulting in the noisy output $\hat{y}(k)$ can be represented as
%
\begin{equation}
\hat{y}(k)=y(k)+\nu(k).
\end{equation}

Note that the model in Eq.~\eqref{eq:tfwithout} is equivalent to the following difference equation:
%
\begin{equation}
\begin{split}
y(k)=-a_{1}y(k-1)-a_{2}y(k-2)-\cdots - a_{N}y(k-N)\\
+b_{0}u(k)+b_{1}u(k-1)+b_{2}u(k-2)+\cdots+b_{M}u(k-M),
\end{split}
\end{equation}

\noindent and consequently $\hat{y}(k)$ can be represented by the following difference equation:
%
\begin{equation}
\label{eq:differencewithnoise}
\begin{split}
\hat{y}(k)=-a_{1}y(k-1)-a_{2}y(k-2)-\cdots - a_{N}y(k-N)\\
+b_{0}u(k)+b_{1}u(k-1)+b_{2}u(k-2)+\cdots+b_{M}u(k-M)+\nu(k).
\end{split}
\end{equation}

The z-transform of Eq.~\eqref{eq:differencewithnoise} is
%
\begin{equation}
\begin{split}
\hat{Y}(z)=(-a_{1}z^{-1}-a_{2}z^{-2}-\cdots-a_{N}z^{-N})Y(z)\\
+(b_{0}+b_{1}z^{-1}+b_{2}z^{-2}+\cdots+b_{M}z^{-M})U(z)+\nu(z),
\end{split}
\end{equation}
dividing both sides by $U(z)$, it is equivalent to:
\begin{equation}
\frac{\hat{Y}(z)}{U(z)}=\hat{G}(z)=[1-A(z)]G(z)+U(z)B(z)+\frac{\nu(z)}{U(z)}.
\end{equation}

Considering $\hat{G}(z)=G(z)+\Delta{G_{Q}(z)}$ and Eq.~\eqref{eq:tfwithout}, the intended expression can be obtained as
%
\begin{equation}
\Delta{G_{Q}(z)}=\frac{\nu(z)}{U(z)} 
\end{equation}
\end{proof}

Eq.~\eqref{eq:deltag_var} indicates that the modeling of quantization noise effect on the plant transfer function demands an estimation of the variance of quantization noise $\sigma_{Q}^{2}$. To obtain such estimation, the methodology presented by Widrow and Koll\'ar~\cite{widrow2008quantization} is used. The authors provide an algorithm for computing the mean square of the output noise due to $Q1$ and $Q2$. Considering that each quantizer contribution can be represented by a white noise with zero mean (Assumption~\ref{whitenoise}), the mean square can be considered as the variance of noise ($\sigma_{Q}^{2}$), which can be calculated as
%
\begin{equation}
\sigma_{Q}^{2}=\frac{\lambda_{1}^{2}\cdot \sum {h_{Q1}^{2}}+\lambda_{2}^{2}\cdot \sum {h_{Q2}^{2}}}{12},
\end{equation}

\noindent where $\lambda_{1}$ and $\lambda_{2}$ are the scale factor for the quantization $Q1$ and $Q2$, respectively; and $\sum {h_{Q1}^{2}}$ and $\sum {h_{Q2}^{2}}$ are the squares sums of the impulse response obtained by means of a transfer function from $Q1$ and $Q2$ to the plant output, respectively.

The transfer functions $H_{Q1}(z)$ and $H_{Q2}(z)$, from which the squares sums of the impulse response are needed, can be obtained through the analysis of the system in Figure~\ref{fig:sampledsystem}. Considering that the quantizer $Q1$ is the source of a white noise with zero mean $\nu_{1}$ and $Q2$ is the source of a white noise with zero mean $\nu_{2}$, the following equation can be written for the system in Figure~\ref{fig:sampledsystem}
%
\begin{equation}
\resizebox{0.48\textwidth}{!}{
$\hat{Y}(z)=\nu_{1}(z)C(z)G(z)+\nu_{2}(z)G(z)+R(z)C(z)G(z)-\hat{Y}(z)C(z)G(z),$
}
\end{equation}

\noindent which can be rewritten as
%
\begin{equation}
\label{eq:outputfunctions}
\resizebox{0.48\textwidth}{!}{
$\hat{Y}(z)=\frac{C(z)G(z)}{1+C(z)G(z)}R(z)+\frac{C(z)G(z)}{1+C(z)G(z)}\nu_{1}(z) \\
+\frac{G(z)}{1+C(z)G(z)}\nu_{2}(z).$
}
\end{equation}

Thus, $H_{Q1}(z)$ and $H_{Q2}(z)$ can be extracted from Eq.~\eqref{eq:outputfunctions}:
\begin{equation}
H_{Q1}(z)=\frac{C(z)G(z)}{1+C(z)G(z)},
\end{equation}
and
\begin{equation}
H_{Q2}(z)=\frac{G(z)}{1+C(z)G(z)}.
\end{equation}

These equations are sufficient for computing $\sigma_{Q}^{2}$. However, there is still a gap for a complete model for $\Delta{G_{Q}(z)}$ in Eq.~\eqref{eq:deltag_var}, {\it i.e.}, the z-transform of control action $U(z)$. This indicates that the quantization noise effect on a discrete plant model mainly depends on three main factors:
%
\begin{itemize}
\item \textbf{Sample time:} the sample time effect is reflected over pulse transfer function that covers the effects of DAC ant its interpolator (ZOH);
\item \textbf{ADC and DAC step: } these are the main component of quantization noise power, but an $8$-bits converter already reduces the effect of quantization noises in the output for negligible values.
\item \textbf{Signal-to-noise ratio (SNR): } the dependence on the value of $U(z)$ shows the importance of the SNR for handling quantization noise; for a high SNR of the digital controller, the quantization effect is negligible. This happens during the transient of the system, when high control actions are employed. However, it can be relevant during the steady state. A digital controller with integral action might be sufficient for eliminating such influence on the steady state.
\end{itemize}

An alternative expression for $\Delta{G_{Q}(z)}$ can be obtained from Eq.~\eqref{eq:deltag_var}, transforming it into a function of the reference $R(z)$. Once $U(z)$ is the control signal obtained via digital controller $C(z)$ and error signal, {\it i.e.},
%
\begin{equation}
U(z)=C(z)\cdot[R(z)-Y(z)],
\end{equation}

\noindent where $Y(z)$ is the first term of Eq.~\eqref{eq:outputfunctions}, thus:
\begin{equation}
\label{eq:controlsignal}
\resizebox{0.48\textwidth}{!}{
$U(z)=C(z)\cdot \left[R(z)-\frac{C(z)G(z)}{1+C(z)G(z)}R(z)\right]=\frac{C(z)}{1+C(z)G(z)}R(z).$
}
\end{equation}

Substituting Eq.~\eqref{eq:controlsignal} in Eq.~\eqref{eq:deltag_var}, a model for $\Delta{G_{Q}(z)}$ depending on $R(z)$ can be obtained:
%
\begin{equation}
\Delta{G_{Q}(z)}=\sigma^{2}_{Q}\cdot \frac{1+C(z)G(z)}{C(z)} \cdot \frac{1}{R(z)},
\end{equation}

\noindent where $R(z)$ can be considered as a step signal (for stabilization purpose) with height equal to $r$, which can be rewritten as
\begin{equation}
\Delta{G_{Q}(z)}=\frac{\sigma^{2}_{Q}}{r}\cdot \frac{1+C(z)G(z)}{C(z)} \cdot \frac{z-1}{z}.
\end{equation}

A more confident description of $\Delta{G_{Q}(z)}$ can be obtained considering FWL effects on digital controller coefficients, as described in Bessa {\it et al.}~\cite{Bessa16}, {\it i.e.}, using $FWL[C(z)]$ instead of the nominal $C(z)$:
%
\begin{equation}
\label{eq:deltaG_final}
\Delta{G_{Q}(z)}=\frac{\sigma^{2}_{Q}}{r}\cdot \frac{1+FWL[C(z)]G(z)}{FWL[C(z)]} \cdot \frac{z-1}{z}.
\end{equation}

Finally, Eq.~\eqref{eq:deltaG_final} is a suitable frequency domain model for quantization noise effects on closed-loop systems. This model depends on the plant pulse transfer function ($G(z)$), which is described by Eq.~\eqref{eq:pulsetf}, the height of step reference signal $r$, and the digital controller model.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Error Bound Verification}

%As described in the literature, computation using FWL leads to rounding and truncation errors~\cite{Oppenheim1999}. In the present work, impairments present on a closed-loop  system output, due to coefficient rounding and arithmetic-operation results, are considered. Let $x$ be a real number and $x^*$ be a floating-point representation of $x$. The absolute error $\epsilon$, due to the rounding of $x$ represented in floating-point as $x^*$, is 
%
%\begin{equation} \label{absolute-error}
%x^* = x + \epsilon,
%\end{equation}

%\noindent and the relative error $\delta$ by
%
%\begin{equation} \label{relative-error}
%x^* = x\left(1+\delta\right).
%\end{equation}

%The verification of output error bound is proposed, where the output of a closed-loop system, implemented with reduced-word-length floating-point representation, is compared with a reference output of the same structures, implemented using real variables.

%Thus, considering that the number of precision bits $l_{d}$ in reference models is higher than in designed models, the quantization error $E_{r}$ is given by
%\begin{equation} \label{errorfixlimits}
%-2^{-l-1} + 2^{-l_{d}-1} \leq E_{r} \leq 2^{-l-1} - 2^{-l_{d}-1}.
%\end{equation}

%The expression above shows that the error, computed with the proposed method, is affected by the precision of the reference model.

%For the reduced floating-point models, the computed values are saturated to the maximum representable number on overflow, or to the minimum on underflow; the same input is applied to both models. The input vector uses non-deterministic values from $\{ 2^{k-1}-2^{-l}, 2^{-l}, 0, -2^{-l}, 2^{k-1} \}$, that is, values for maximum and minimum amplitudes of the input signal. So, errors due to quantization and saturation are explored.

%During closed-loop operation unrolling, the cumulative error can increase beyond the quantization-error interval. The following literal is then generated, in order to check whether the output error is in accordance with an acceptable margin:
%\begin{equation} \label{errorcheck}
%	l_{no\_error} \Leftrightarrow \arrowvert y - y_{d} \arrowvert \leq M \cdot 2^{-l-1},
%\end{equation}

%\noindent where $y$ is the output of the designed system, $y_{d}$ is the output of the double precision system, and $M~2^{-l-1}$ is the tolerance margin. The verification process searches for the negation of this literal. When such a counterexample is found, it indicates that the output error is higher than expected for that controller realization.
